---
title: "MBTA EDA"
author: "Hao"
date: "2022-12-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE)
```
```{r}
library(dplyr)
library(ggplot2)
library(fmsb)
```


# Import dataset

The data, download from https://cdn.mbta.com/archive/archived_feeds.txt, includes date: 20211111, 20211209, 20220121, 20220204, 20220304, 20220426, 20220502, 20220603, 20220706, 20220803, 20220916, 20221013.For convennience, I combined 12-calender files into one, calender.txt, and also for routes.
```{r}
calendar<-read.csv("calendar.txt")
routes<-read.csv("routes.txt")
```

```{r}
frequence<-count(calendar,service_schedule_typicality)
barplot(height=frequence$n,
        names=frequence$service_schedule_typicality,
        col=c("bisque","lightblue","coral","aquamarine2"),
        xlab="service_schedule tyoicality",
        ylab="Count")
```
This graph shows the frequency of each type.


```{r}
par(mfrow=c(2,2))
type_1<-calendar%>%filter(service_schedule_typicality==1)
f_1<-count(type_1,rating_description)
barplot(height = f_1$n,names=f_1$rating_description,main = "type 1",col=c("bisque","lightblue","coral","aquamarine2"))
type_3<-calendar%>%filter(service_schedule_typicality==3)
f_2<-count(type_3,rating_description)
barplot(height = f_2$n,names=f_2$rating_description,main = "type 3",col=c("bisque","lightblue","coral","aquamarine2"))
type_4<-calendar%>%filter(service_schedule_typicality==4)
f_3<-count(type_4,rating_description)
barplot(height = f_3$n,names=f_3$rating_description,main = "type 4",col=c("bisque","lightblue","coral","aquamarine2"))
type_5<-calendar%>%filter(service_schedule_typicality==5)
f_4<-count(type_5,rating_description)
barplot(height = f_4$n,names=f_4$rating_description,main = "type 5",col=c("bisque","lightblue","coral","aquamarine2"))
```
This graph shows the frequency of different seasons for each type.

```{r}
route_id<-c("Red","Orange","Blue","Green-B","Green-C","Green-D","Green-E")
color<-c("burlywood1","cadetblue1","darkkhaki","darkslategray1","darkseagreen4","darkolivegreen1")
par(mfrow=c(2,2))
type_1<-routes[routes$route_pattern_typicality==1,]
type_1<-type_1[type_1$route_id%in%route_id,]
r_1<-count(type_1,route_id)
barplot(height = r_1$n,names=r_1$route_id,main = "type 1",col=color)


type_2<-routes[routes$route_pattern_typicality==2,]
type_2<-type_2[type_2$route_id%in%route_id,]
r_2<-count(type_2,route_id)
barplot(height = r_2$n,names=r_2$route_id,main = "type 2",col=color,xlim = c(0,50),ylim=c(0,50))


type_3<-routes[routes$route_pattern_typicality==3,]
type_3<-type_3[type_3$route_id%in%route_id,]
r_3<-count(type_3,route_id)
barplot(height = r_3$n,names=r_3$route_id,main = "type 3",col=color)


type_4<-routes[routes$route_pattern_typicality==4,]
type_4<-type_4[type_4$route_id%in%route_id,]
r_4<-count(type_4,route_id)
barplot(height = r_4$n,names=r_4$route_id,main = "type 4",col=color)
```
This graph shows the frequencies of the different routes of each type. I only select some common lines, including "Red","Orange","Blue","Green-B","Green-C","Green-D","Green-E".



